### 2018-10-15 Chapters/Chapter2.tex
com

\begin{align}
\label{PertNoEMF}
&\bra{\phi'} -\frac{\hat{H_t^2}}{\text{U}} \ket{\phi} = \nonumber \\
&-\frac{1}{\text{U}} \bra{\phi'} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2} \sum_{\langle i',j' \rangle, 					\sigma_3, \sigma_4} (t^2 \delta_{\sigma_1 \sigma_2}\delta_{\sigma_3 \sigma_4} + t(\delta_{\sigma_1 \sigma_2}		\boldsymbol{\Delta}_{i'j'}\boldsymbol{\sigma}_{\sigma_3 \sigma_4} + \delta_{\sigma_3 \sigma_4}						\boldsymbol{\Delta}_{ij}\boldsymbol{\sigma}_{\sigma_1 \sigma_2}) + \mathcal{O}(\boldsymbol{\Delta}^2)) 				\hat{c}_{i\sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{i' \sigma_3}^\dagger \hat{c}_{j'\sigma_4}					\ket{\phi} = \nonumber \\
&-\frac{1}{\text{U}} \bra{\phi'} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} (t^2 						\delta_{\sigma_1 \sigma_2}\delta_{\sigma_3 \sigma_4} + t\boldsymbol{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4}			\boldsymbol{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \boldsymbol{\sigma}_{\sigma_3 					\sigma_4}) + \mathcal{O}(\boldsymbol{\Delta}^2)) \hat{c}_{i\sigma_1}^\dagger \hat{c}_{i\sigma_4} \hat{c}_{j 		\sigma_2} \hat{c}_{j \sigma_3}^\dagger \ket{\phi} = \nonumber \\
&-\frac{1}{\text{U}} \bra{\phi'} \hat{H}_J + \hat{H}_D + \mathcal{O}(\boldsymbol{\Delta}^2))\ket{\phi}
\end{align}

In the second line we imposed $i=j'$ and $j=i'$ since that's the only way $\hat{H_t^2}\ket{\phi}$ has no double occupied sites and so the matrix element does not vanish. We also used $\bs{\Delta}_{ij} = - \bs{\Delta}_{ji}$. We will neglect the $\mathcal{O}(\boldsymbol{\Delta}^2)$ term since $t$ is dominant. We defined:

\begin{align}
\hat{H}_J &= -\frac{t^2}{\text{U}} \sum_{\langle i,j \rangle, \sigma, \sigma'} \hat{c}_{i\sigma}^\dagger \hat{c}_{i\sigma'} \hat{c}_{j\sigma} \hat{c}_{j \sigma'}^\dagger \\
\hat{H}_D &= -\frac{t}{\text{U}}\sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} \boldsymbol{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4}	\boldsymbol{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \boldsymbol{\sigma}_{\sigma_3 	\sigma_4}) \hat{c}_{i\sigma_1}^\dagger \hat{c}_{i\sigma_4} \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger
\end{align}

@@

These two relations are important and will be used in the next chapter:

\begin{align}
&\sum_{\sigma \sigma'} \left(\frac{1}{2}\delta_{\sigma \sigma'} + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma' \sigma}\right)\left(\frac{1}{2}\delta_{\sigma \sigma'}-\boldsymbol{S}_j\boldsymbol{\sigma}_{\sigma \sigma'}\right) =\frac{1}{2} - \sum_{ab} S_i^aS_j^b	2\delta_{ab} = \frac{1}{2} - 2\boldsymbol{S}_i\boldsymbol{S}_j \label{SpinRel1} \\
&\sum_{\sigma_1, \sigma_2, \sigma_3, \sigma_4}\boldsymbol{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4}\boldsymbol{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \boldsymbol{\sigma}_{\sigma_3\sigma_4})\left(\frac{1}{2}\delta_{\sigma_1 \sigma_4} + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma_4 \sigma_1}\right)\left(\frac{1}{2}\delta_{\sigma_2 \sigma_3}-\boldsymbol{S}_j\boldsymbol{\sigma}_{\sigma_2 \sigma_3}\right)= -4i \boldsymbol{\Delta}_{ij} \boldsymbol{S}_i \times \boldsymbol{S}_j \label{SpinRel2}
\end{align}

@@

Where $\boldsymbol{D}_{ij} = \frac{4it}{\text{U}}\boldsymbol{\Delta}_{ij}$. The total effective Hamiltonian is:

>>>>

\begin{align}
\label{PertNoEMF}
&\bra{\phi'} -\frac{\hat{H_t^2}}{\text{U}} \ket{\phi} = \nonumber \\
&= -\frac{1}{\text{U}} \bra{\phi'}\sum_{\langle i,j \rangle, \sigma_1, \sigma_2} \sum_{\langle i',j' \rangle, 					\sigma_3, \sigma_4} \left( \delta_{\sigma_1, \sigma_2} t + \bs{\Delta}_{ij} \bs{\sigma}_{\sigma_1, \sigma_2} \right) \left( \delta_{\sigma_3, \sigma_4} t + \bs{\Delta}_{i'j'} \bs{\sigma}_{\sigma_3, \sigma_4} \right)\hat{c}_{i\sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{i' \sigma_3}^\dagger \hat{c}_{j'\sigma_4} \ket{\phi} = \nonumber \\
&= -\frac{1}{\text{U}} \bra{\phi'}\sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} \left( \delta_{\sigma_1, \sigma_2} t + \bs{\Delta}_{ij} \bs{\sigma}_{\sigma_1, \sigma_2} \right) \left( \delta_{\sigma_3, \sigma_4} t - \bs{\Delta}_{ij} \bs{\sigma}_{\sigma_3, \sigma_4} \right)\hat{c}_{i\sigma_1}^\dagger \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \hat{c}_{i\sigma_4} \ket{\phi} = \nonumber \\
&-\frac{1}{\text{U}} \bra{\phi'} \sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} (t^2 						\delta_{\sigma_1 \sigma_2}\delta_{\sigma_3 \sigma_4} + t\bs{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4}			\bs{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \bs{\sigma}_{\sigma_3 \sigma_4}) - (\bs{\Delta}_{ij}\bs{\sigma}_{\sigma_1 \sigma_2})(\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_3 \sigma_4}) ) \hat{c}_{i\sigma_1}^\dagger \hat{c}_{i\sigma_4} \hat{c}_{j 		\sigma_2} \hat{c}_{j \sigma_3}^\dagger \ket{\phi} = \nonumber \\
& \bra{\phi'} \hat{H}_J + \hat{H}_D + \hat{H}_P \ket{\phi}
\end{align}

In the second line we imposed $i=j'$ and $j=i'$ since that's the only way $\hat{H_t^2}\ket{\phi}$ has no double occupied sites and so the matrix element does not vanish. We also used $\bs{\Delta}_{ij} = - \bs{\Delta}_{ji}$. We defined:

\begin{align}
\hat{H}_J &= -\frac{t^2}{\text{U}} \sum_{\langle i,j \rangle, \sigma, \sigma'} \hat{c}_{i\sigma}^\dagger \hat{c}_{i\sigma'} \hat{c}_{j\sigma} \hat{c}_{j \sigma'}^\dagger \\
\hat{H}_D &= -\frac{t}{\text{U}}\sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} \bs{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4}	\bs{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \bs{\sigma}_{\sigma_3 	\sigma_4}) \hat{c}_{i\sigma_1}^\dagger \hat{c}_{i\sigma_4} \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger \\
\hat{H}_P &= \frac{1}{\text{U}}\sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} (\bs{\Delta}_{ij}\bs{\sigma}_{\sigma_1 \sigma_2})(\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_3 \sigma_4}) \hat{c}_{i\sigma_1}^\dagger \hat{c}_{i\sigma_4} \hat{c}_{j \sigma_2} \hat{c}_{j \sigma_3}^\dagger
\end{align}

@@

Also:

\begin{align*}
&\sum_{\sigma_1, \sigma_2, \sigma_3, \sigma_4} (\bs{\Delta}_{ij}\bs{\sigma}_{\sigma_1 \sigma_2})(\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_3 \sigma_4})\bs{S}_i\bs{\sigma}_{\sigma_4 \sigma_1}\left(-\bs{S}_j\bs{\sigma}_{\sigma_2 \sigma_3}\right) = \\
&= \dots = \\
&= \begin{pmatrix}
S_i^x & S_i^y & S_i^z
\end{pmatrix}
\\
&\begin{pmatrix}
-2(\Delta_{ij}^z)^2 - 2(\Delta_{ij}^y)^2 + 2(\Delta_{ij}^x)^2 & 4\Delta_{ij}^x\Delta_{ij}^y & 4\Delta_{ij}^x\Delta_{ij}^z \\ 4\Delta_{ij}^y\Delta_{ij}^x & -2(\Delta_{ij}^x)^2+2(\Delta_{ij}^y)^2 -2(\Delta_{ij}^z)^2 & 4\Delta_{ij}^y\Delta_{ij}^z \\ 4\Delta_{ij}^z\Delta_{ij}^x & 4\Delta_{ij}^z\Delta_{ij}^y & -2(\Delta_{ij}^x)^2-2(\Delta_{ij}^y)^2+2(\Delta_{ij}^z)^2
\end{pmatrix}
\\
&\begin{pmatrix}
S_j^x \\ S_j^y \\ S_j^z
\end{pmatrix} = \\
&\sum_{\alpha, \beta} S_i^\alpha (\delta_{\alpha \beta} 2\bs{\Delta}_{ij}^2 - 4\Delta_{ij}^\alpha\Delta_{ij}^\beta ) S_j^\beta
\end{align*}

These three relations are important and will be used in the next chapter:

\begin{align}
&\sum_{\sigma \sigma'} \left(\frac{1}{2}\delta_{\sigma \sigma'} + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma' \sigma}\right)\left(\frac{1}{2}\delta_{\sigma \sigma'}-\boldsymbol{S}_j\boldsymbol{\sigma}_{\sigma \sigma'}\right) =\frac{1}{2} - \sum_{ab} S_i^aS_j^b	2\delta_{ab} = \frac{1}{2} - 2\boldsymbol{S}_i\boldsymbol{S}_j \label{SpinRel1} \\
&\sum_{\sigma_1, \sigma_2, \sigma_3, \sigma_4}\boldsymbol{\Delta}_{ij}(\delta_{\sigma_3 \sigma_4}\boldsymbol{\sigma}_{\sigma_1 \sigma_2} - \delta_{\sigma_1 \sigma_2} \boldsymbol{\sigma}_{\sigma_3\sigma_4})\left(\frac{1}{2}\delta_{\sigma_1 \sigma_4} + \boldsymbol{S}_i\boldsymbol{\sigma}_{\sigma_4 \sigma_1}\right)\left(\frac{1}{2}\delta_{\sigma_2 \sigma_3}-\boldsymbol{S}_j\boldsymbol{\sigma}_{\sigma_2 \sigma_3}\right)= -4i \boldsymbol{\Delta}_{ij} \boldsymbol{S}_i \times \boldsymbol{S}_j \label{SpinRel2} \\
& \sum_{\sigma_1, \sigma_2, \sigma_3, \sigma_4} (\bs{\Delta}_{ij}\bs{\sigma}_{\sigma_1 \sigma_2})(\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_3 \sigma_4})\bs{S}_i\bs{\sigma}_{\sigma_4 \sigma_1}\left(-\bs{S}_j\bs{\sigma}_{\sigma_2 \sigma_3}\right) = \sum_{\alpha, \beta} S_i^\alpha (\delta_{\alpha \beta} 2\bs{\Delta}_{ij}^2 - 4\Delta_{ij}^\alpha\Delta_{ij}^\beta ) S_j^\beta \label{SpinRel3}
\end{align}

@@

Where $\boldsymbol{D}_{ij} = \frac{4it}{\text{U}}\boldsymbol{\Delta}_{ij}$. Lastly, using \ref{SpinRel3}, the term proportional to $\bs{\Delta}_{ij}^2$ gives:

\begin{align*}
\hat{H}_P &= \frac{1}{\text{U}}\sum_{\langle i,j \rangle, \sigma_1, \sigma_2, \sigma_3, \sigma_4} (\bs{\Delta}_{ij}\bs{\sigma}_{\sigma_1 \sigma_2})(\bs{\Delta}_{ij} \bs{\sigma}_{\sigma_3 \sigma_4}) \left(\frac{1}{2}\delta_{\sigma_1 \sigma_4} + \bs{S}_i\bs{\sigma}_{\sigma_4 \sigma_1}\right)\left(\frac{1}{2}\delta_{\sigma_2 \sigma_3}-\bs{S}_j\bs{\sigma}_{\sigma_2 \sigma_3}\right) = \\
&\frac{1}{\text{U}} \sum_{\langle i,j \rangle} \sum_{\alpha, \beta} S_i^\alpha (\delta_{\alpha \beta} 2\bs{\Delta}_{ij}^2 - 4\Delta_{ij}^\alpha\Delta_{ij}^\beta ) S_j^\beta = \sum_{\langle i,j \rangle} \bs{S}_i \bs{\Gamma}_{ij} \bs{S}_j
\end{align*}

With $\bs{\Gamma_{ij}}^{\alpha \beta} = \frac{1}{\text{U}}(\delta_{\alpha \beta} 2\bs{\Delta}_{ij}^2 - 4\Delta_{ij}^\alpha\Delta_{ij}^\beta)$.

The total effective Hamiltonian is:

### 2018-10-15 Chapters/Chapter3.tex
com

Where $\hat{D}$ is the double occupancy number operator and $\hat{T}(t) = \sum_{\langle i,j \rangle, \sigma} t_{ij}(t)\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma}$ is the hopping operator. Let $e\bs{R}_{ij}\vec{A} = \alpha_{ij} e^{i \theta_{ij}}$, with $\alpha_{ij}$ a real number, then we can apply the Jacobi–Anger expansion:

@@

Now, $\alpha_{ij} = \text{Re}(e\bs{R}_{ij}\vec{A})$. For a square lattices aligned with the coordinate system so that $\vec{a}_1=\hat{e}_x$ and $\vec{a}_2=\hat{e}_y$, we will have $\bs{R}_{ij} = \pm a\hat{e}_x,\pm a\hat{e}_y$, where $a$ is the lattice constant. Using $\vec{A}=\frac{iE_0}{\omega\sqrt{1+\lambda^2}}(\hat{e}_x+i\lambda\hat{e}_y)$ we have:

@@

@@

\begin{align*}
\hat{H}'(\bs{r}) \psi'_{\bs{R}} (\bs{r}) &= \left[ \frac{(\bs{p}-e\bs{A})^2}{2m} +U(\bs{r}) \right] e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \psi_{\bs{R}}(\bs{r}) = \\
&= e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \left[ \frac{\bs{p}^2}{2m} + U(\bs{r}) \right] \psi_{\bs{R}}(\bs{r}) = e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \hat{H}(\bs{r}) \psi_{\bs{R}}(\bs{r})
\end{align*}

>>>>

Where $\hat{D}$ is the double occupancy number operator and $\hat{T}(t) = \sum_{\langle i,j \rangle, \sigma} t_{ij}(t)\hat{c}_{i \sigma}^\dagger \hat{c}_{j \sigma}$ is the hopping operator. Let $e\bs{R}_{ij}\vec{A} = \alpha_{ij} e^{i \theta_{ij}}$, with $\alpha_{ij} = \pm|e\bs{R}_{ij}\vec{A}|$ in such a way that:

\begin{align}
\alpha_{ij} &= -\alpha_{ji} \label{alphaSym} \\
\theta_{ij} &= \theta_{ji} \label{thetaSym} \\
\end{align}

and $\theta_{ij} \in \left[0,\pi\right)$. Then we can apply the Jacobi–Anger expansion:

@@

Now, $\alpha_{ij} = \pm|e\bs{R}_{ij}\vec{A}|$. For a square lattices aligned with the coordinate system so that $\vec{a}_1=\hat{e}_x$ and $\vec{a}_2=\hat{e}_y$, we will have $\bs{R}_{ij} = \pm a\hat{e}_x,\pm a\hat{e}_y$, where $a$ is the lattice constant. Using $\vec{A}=\frac{iE_0}{\omega\sqrt{1+\lambda^2}}(\hat{e}_x+i\lambda\hat{e}_y)$ we have:

@@

\begin{subsection}{Honeycomb lattice}

Now, in \ref{Jij2}, consider that the $\alpha_{ij}$ values are given by $\alpha_{ij} = \pm|e\bs{R}_{ij}\vec{A}|$ where $\bs{R}_{ij}$ are the displacement vectors in a honeycomb lattice. There are six such vectors:

\begin{align}
\bs{R}_1^\pm &= \pm a\hat{e}_x \\
\bs{R}_2^\pm &= \pm a\frac{1}{2}(\hat{e}_x + \sqrt{3}\hat{e}_y) \\
\bs{R}_3^\pm &= \pm a\frac{1}{2}(\hat{e}_x - \sqrt{3}\hat{e}_y)
\end{align}

These six possible directions will lead to $\alpha_a^\pm = \pm |e\bs{R}_a\vec{A}|$ where $\vec{A}=\frac{iE_0}{\omega\sqrt{1+\lambda^2}}(\hat{e}_x+i\lambda\hat{e}_y)$:

\begin{align}
\alpha_1^\pm &= \pm \frac{eaE_0}{\omega\sqrt{1+\lambda^2}} \\
\alpha_2^\pm &= \alpha_3^\pm = \pm \frac{eaE_0}{2\omega} \sqrt{\frac{1+3\lambda^2}{1+\lambda^2}} = \pm\frac{eaE_0}{2\omega}\sqrt{1+\frac{2\lambda^2}{1+\lambda^2}}
\end{align}

We see that again, this result will not depend on the helicity of the applied electric field.

\end{subsection}

@@

\begin{align*}
\hat{H}'(\bs{r}) \psi'_{\bs{R}} (\bs{r}) &= \left[ \frac{(\bs{p}-e\bs{A})^2}{2m} +U(\bs{r}) \right] e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \psi_{\bs{R}}(\bs{r}) = \\
&=e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \left[ \frac{(\bs{p} - i\hbar \bs{\nabla_r}(i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)) -e\bs{A})^2}{2m} +U(\bs{r}) \right] = \\
&= e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \left[ \frac{\bs{p}^2}{2m} + U(\bs{r}) \right] \psi_{\bs{R}}(\bs{r}) = e^{i\frac{e}{\hbar}\int_{\bs{R}}^{\bs{r}} d\bs{r}'\bs{A}(\bs{r}',t)} \hat{H}(\bs{r}) \psi_{\bs{R}}(\bs{r})
\end{align*}
